---
title: "Yolk Data Analysis"
author: "Naomi Laven"
date: "2023-10-19"
output: html_document
editor_options: 
  chunk_output_type: console
---
### Packages
```{r}
pacman::p_load(ggplot2, tidyverse, car, patchwork)
```

### Data Load-In and transformation

```{r}
#NEW RESULTS WITH all data readjusted as in supplementary section 2.

yolkdatv3<-read.csv("Datafiles/yolkresultsv3.csv")
#Make momid a factor not a number
yolkdatv3$MomID<- factor(yolkdatv3$MomID,)

yolkdatv3$logE <- log(yolkdatv3$finalE.pg.mg.)
yolkdatv3$logT <- log(yolkdatv3$finalT.pg.mg.)

```


### YOlk Mass Associations

```{r}
mass_assoc_T <- ggplot(yolkdatv3, aes(x = EggMass, y = logT, color = MomID)) + 
  geom_point(aes(shape = MomGeno), size = 6) + 
  theme_classic() + 
  geom_smooth(method = lm, color = "black") +
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  labs(color = "Mother's ID") + 
  xlab("Egg Mass (g)") + 
  ylab("Log Testosterone Concentration (pg/mg)")

mass_assoc_T

lm_mass_assoc_T <- summary(lm(yolkdatv3$EggMass~yolkdatv3$logT))
lm_mass_assoc_T
```

```{r}
mass_assoc_E <- ggplot(yolkdatv3, aes(x = EggMass, y = logE, color = MomID)) + 
  geom_point(aes(shape = MomGeno), size = 6) + 
  theme_classic() + 
  geom_smooth(method = lm, color = "black") +
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  labs(color = "Mother's ID") + 
  xlab("Egg Mass (g)") + 
  ylab("Log Estradiol Concentration (pg/mg)")

mass_assoc_E

lm_mass_assoc_E <- summary(lm(yolkdatv3$EggMass~yolkdatv3$logE))
lm_mass_assoc_E
```

```{r}
combinedMass <- mass_assoc_T / mass_assoc_E
combinedMass

ggsave(filename = "yolkhorMass_assoc.png", path = "Pogona Hormones/output", width = 15, height = 20)
```

Laying order?
```{r}
ggplot(yolkdatv3, aes(y = EggMass, x = EggNo)) + geom_point(aes(shape = MomGeno, color = MomID), size = 6) + geom_smooth(method = "lm", color = "black")+ theme_classic() +
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) + ylim(2.5, 5) + xlim(1, 25) + ylab("Egg Mass (g)") + xlab("Egg Number")

ggsave(filename = "Layorder_vs_Mass.png", path = "Pogona Hormones/output", width = 10, height = 20)

```

```{r}
plot(yolkdatv3$EggNo~yolkdatv3$logE)
summary(lm(yolkdatv3$EggNo~yolkdatv3$logE))

```

```{r}
plot(yolkdatv3$EggNo~yolkdatv3$logT)
summary(lm(yolkdatv3$EggNo~yolkdatv3$logT))

```


### Corrected for Yolk Mass?

```{r}
yolkdatv3$masscorrectedE <- (yolkdatv3$finalE.pg.mg.*yolkdatv3$EggMass)
yolkdatv3$masscorrectedT <- (yolkdatv3$finalT.pg.mg.*yolkdatv3$EggMass)
```

```{r}
qqPlot(yolkdatv3$mass_corrected_E.pg.mg.)
```
```{r}
qqPlot(yolkdatv3$mass_corrected_T.pg.mg.)
```

```{r}
yolkdatv3$logmcE <- (log(yolkdatv3$masscorrectedE))
yolkdatv3$logmcT <- (log(yolkdatv3$masscorrectedT))
```

```{r}
qqPlot(yolkdatv3$logmcE)
```
```{r}
qqPlot(yolkdatv3$logmcT)
```

```{r}

ggplot(yolkdatv3, aes(x = logmcE, y = logmcT, color = MomID)) + 
  geom_point(aes(shape = MomGeno), size = 6) + 
  theme_classic() + 
  geom_smooth(method = lm, color = "black") +
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  labs(color = "Mother's ID") + 
  xlab("Log Testosterone Concentration (total)") + 
  ylab("Log Estradiol Concentration (total)")


summary(lm(yolkdatv3$logmcE~yolkdatv3$logmcT))

ggsave(filename = "MASSCORRyolkhor_assoc.png", path = "Pogona Hormones/output", width = 16, height = 15)
```

```{r}
mass_assoc_T <- ggplot(yolkdatv3, aes(x = EggMass, y = logmcT, color = MomID)) + 
  geom_point(aes(shape = MomGeno), size = 6) + 
  theme_classic() + 
  geom_smooth(method = lm, color = "black") +
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  labs(color = "Mother's ID") + 
  xlab("Egg Mass (grams)") + 
  ylab("Log Testosterone Concentration (total)")

mass_assoc_T

lm_mass_assoc_T <- summary(lm(yolkdatv3$EggMass~yolkdatv3$logmcT))
lm_mass_assoc_T
```

```{r}
mass_assoc_E <- ggplot(yolkdatv3, aes(x = EggMass, y = logmcE, color = MomID)) + 
  geom_point(aes(shape = MomGeno), size = 6) + 
  theme_classic() + 
  geom_smooth(method = lm, color = "black") +
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  labs(color = "Mother's ID") + 
  xlab("Egg Mass (grams)") + 
  ylab("Log Estradiol Concentration (total)")

mass_assoc_E

lm_mass_assoc_E <- summary(lm(yolkdatv3$EggMass~yolkdatv3$logmcE))
lm_mass_assoc_E
```

```{r}
combinedMass <- mass_assoc_T / mass_assoc_E
combinedMass

ggsave(filename = "yolkhorMass_assoc.png", path = "Pogona Hormones/output", width = 13, height = 15)
```


### T/E2 Concentration Correlation Plot
```{r}
# Create the plot
hormoneassoc <- ggplot(yolkdatv3, aes(x = logE, y = logT, color = MomID)) + 
  geom_point(aes(shape = MomGeno), size = 6) + 
  theme_classic() + 
  geom_smooth(method = lm, color = "black") +
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  labs(color = "Mother's ID") + 
  xlab("Log Testosterone Concentration (pg/mg)") + 
  ylab("Log Estradiol Concentration (pg/mg)")

# Print the plot

print(hormoneassoc)
ggsave(filename = "yolkhor_assoc.png", path = "Pogona Hormones/output", width = 16, height = 15)
```


### Clutch Averages Calcuation
```{r}
avg_concentrations <- yolkdatv3 %>%
  group_by(MomID) %>%
  summarize(Avg_T = mean(logT),
            Avg_E2 = mean(logE), AvgEggMass = mean(EggMass))

# Print the table
print(avg_concentrations)


```

```{r}
#For Mass Corrected Data (Wed 17th July)
clutch_avgs <- yolkdatv3 %>%
  group_by(MomID) %>%
  summarize(Avg_mcT = mean(logmcT),
            Avg_mcE2 = mean(logmcE), AvgEggMass = mean(EggMass))

# Print the table
print(clutch_avgs)


```

```{r}
#Quickly checking if correlation between T and E persists for weight-corrected clutch averages
plot(clutch_avgs$Avg_mcT~clutch_avgs$Avg_mcE2)
```


### Visualising the hormones of clutch mothers

```{r}
pogdat<-read.csv("Datafiles/Final_AdultHormone_DataSheet.csv")

#Set factors and numerical to avoid issues
pogdat$Gen_Sex<-as.factor(pogdat$Gen_Sex)
pogdat$Sex<-as.factor(pogdat$Sex)
pogdat$Testosterone<-as.numeric(pogdat$Testosterone)

#Make sure collection periods are presented in order
pogdat$Col_Period<- factor(pogdat$Col_Period, levels=c('Earl', 'Mid', 'Late'))

```

```{r}
#Refine dataset down to mothers of tested clutches only 

targetIDs<-c("3505", "9152", "9520", "9555", "9559", "9570", "81803", "82142")
pogdat_hormoms<- 
  pogdat %>%
  filter(DragonID %in% targetIDs)

pogdat_hormoms$DragonID<- factor(pogdat_hormoms$DragonID, levels=c("3505", "9152", "9520", "9555", "9559", "9570", "81803", "82142"))

#Remove CVs over 50
pogdat_hormoms <- pogdat_hormoms %>%
  filter(Testosterone_CV <= 50)

```

```{r}
#log transform hromone data 

pogdat_hormoms$logT<-log(pogdat_hormoms$Testosterone)
pogdat_hormoms$logE<-log(pogdat_hormoms$Estrogen_Final)

```

```{r}

#Estradiol of Yolk Mothers

Emother <- ggplot(pogdat_hormoms, aes(x = DragonID, y = logE, fill = Gen_Sex)) + geom_boxplot() + geom_point(aes(shape = Col_Period), size = 6, alpha = 0.75) + theme_classic() + scale_fill_manual(values = c("#809063", "#dab0da"))+
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  labs(fill = "Genotype") + 
  xlab("Dragon ID") + 
  ylab("Log Estradiol Concentration (pg/ml)")

Emother
```

```{r}

#Testosterone of Yolk Mothers

Tmother<-ggplot(pogdat_hormoms, aes(x = DragonID, y = logT, fill = Gen_Sex)) + geom_boxplot() + geom_point(aes(shape = Col_Period), size = 6, alpha = 0.75) + theme_classic() + scale_fill_manual(values = c("#809063", "#dab0da"))+
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  labs(fill = "Genotype") +  labs(shape = "Sample Period")+
  xlab("Dragon ID") + 
  ylab("Log Testosterone Concentration (pg/ml)")

Tmother
```

```{r}
motherhormones <- Tmother / Emother
motherhormones
ggsave(filename = "motherhormones.png", path = "Pogona Hormones/output", width = 16, height = 15)
```

#### Mother's Average Hormone Concentrations 

```{r}
avg_Mother <- pogdat_hormoms %>%
  group_by(DragonID) %>%
  summarize(Avg_T = mean(logT),
            Avg_E2 = mean(logE))

# Print the table
print(avg_Mother)

```

#### Combine mother avergaes and offspring individual yolks
```{r}
avg_Mother <- avg_Mother %>%
  rename(MomID = DragonID)
combined_data <- merge(yolkdatv3, avg_Mother, by = "MomID")

```

```{r}
#Visualise

#Testosterone 
ggplot(combined_data, aes(x = Avg_T, y = logT, color = MomID)) + 
  geom_point(aes(shape = MomGeno),size = 6) + 
  theme_classic() + 
  geom_smooth(method = lm, color = "black") +
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  labs(color = "Mother's ID") + 
  xlab("Log Testosterone Concentration (pg/ml)") + 
  ylab("Log Testosterone Concentration (pg/mg)")+
  scale_shape_manual(values = shape_mapping)

summary(lm(combined_data$Avg_T~combined_data$logT))

```

```{r}
#Estradiol
ggplot(combined_data, aes(x = Avg_E2, y = logE, color = MomID)) + 
  geom_point(aes(shape = MomGeno),size = 6) + 
  theme_classic() + 
  geom_smooth(method = lm, color = "black") +
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  labs(color = "Mother's ID") + 
  xlab("Log Testosterone Concentration (pg/ml)") + 
  ylab("Log Testosterone Concentration (pg/mg)")+
  scale_shape_manual(values = shape_mapping)

summary(lm(combined_data$Avg_E2~combined_data$logE))

ggsave(filename = "motheroffspringE2_indv.png", path = "Pogona Hormones/output", width = 13, height = 15)
```



#### Combine mother/offspring averages
```{r}
avg_Mother <- avg_Mother %>%
  rename(MomID = DragonID)

mother_offspring_avghormones <- avg_concentrations %>%
  left_join(avg_Mother, by = "MomID")


```


```{r}
mc_motheroffspring <- clutch_avgs %>%
  left_join(avg_Mother, by = "MomID")
```



#### Add genotype info back in for mums

```{r}
mom_geno_values <- c("ZW", "ZW", "ZZ", "ZZ", "ZZ", "ZW", "ZZ", "ZW")
mother_offspring_avghormones <- mother_offspring_avghormones %>%
  mutate(MomGeno = mom_geno_values)
```

```{r}
mom_geno_values <- c("ZW", "ZW", "ZZ", "ZZ", "ZZ", "ZW", "ZZ", "ZW")
mc_motheroffspring <- mc_motheroffspring %>%
  mutate(MomGeno = mom_geno_values)
```


### Is clutch hromones associated with clutch mean egg mass?

```{r}
#Testosterone
T_eggmass <- ggplot(yolkdatv3, aes(x = AvgEggMass, y = Avg_T.x, color = MomID)) + 
  geom_point(aes(shape = MomGeno),size = 6) + 
  theme_classic() + 
  geom_smooth(method = lm, color = "black") +
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  labs(color = "Mother's ID") + 
  xlab("Mean Egg Mass (grams)") + 
  ylab("Mean Log Testosterone Concentration (pg/ml)")+
  scale_shape_manual(values = shape_mapping)

summary(lm(mother_offspring_avghormones$AvgEggMass~mother_offspring_avghormones$Avg_T.x))

```

```{r}
#Estradiol
E_eggmass <- ggplot(mother_offspring_avghormones, aes(x = AvgEggMass, y = Avg_E2.x, color = MomID)) + 
  geom_point(aes(shape = MomGeno),size = 6) + 
  theme_classic() + 
  geom_smooth(method = lm, color = "black") +
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  labs(color = "Mother's ID") + 
  xlab("Mean Egg Mass (grams)") + 
  ylab("Mean Log Estradiol Concentration (pg/ml)")+
  scale_shape_manual(values = shape_mapping)

  summary(lm(mother_offspring_avghormones$AvgEggMass~mother_offspring_avghormones$Avg_E2.x))

```


```{r}
eggmass_hormones <- T_eggmass/E_eggmass
eggmass_hormones
ggsave(filename = "eggmass_hormones.png", path = "Pogona Hormones/output", width = 16, height = 15)
```


#### Is mother's estradiol related to egg size?

```{r}

ggplot(mc_motheroffspring, aes(x = Avg_E2, y = AvgEggMass, color = MomID)) + 
  geom_point(aes(shape = MomGeno),size = 6) + 
  theme_classic() + 
  geom_smooth(method = lm, color = "black") +
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  labs(color = "Mother's ID") + 
  xlab("Log Estradiol Concentration (pg/ml)") + 
  ylab("Mean Egg Mass (grams)")+
  scale_shape_manual(values = shape_mapping)

  summary(lm(mc_motheroffspring$AvgEggMass~mc_motheroffspring$Avg_E2))


ggsave(filename = "motherE_vs_Mass.png", path = "Pogona Hormones/output", width = 13, height = 15)
```



#### Regressions
```{r}
shape_mapping <- c("ZW" = 17, "ZZ" = 16)

```

```{r}

momoff_T <- ggplot(mother_offspring_avghormones, aes(x = Avg_T.y, y = Avg_T.x, color = MomID)) + 
  geom_point(aes(shape = MomGeno),size = 6) + 
  theme_classic() + 
  geom_smooth(method = lm, color = "black") +
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  labs(color = "Mother's ID") + 
  xlab("Log Testosterone Concentration (pg/ml)") + 
  ylab("Log Testosterone Concentration (pg/mg)")+
  scale_shape_manual(values = shape_mapping)

momoff_T

quickmodelT <- summary(lm(mother_offspring_avghormones$Avg_T.x~mother_offspring_avghormones$Avg_T.y))

quickmodelT 
```

```{r}

momoff_E <- ggplot(mother_offspring_avghormones, aes(x = Avg_E2.y, y = Avg_E2.x, color = MomID)) + 
  geom_point(aes(shape = MomGeno), size = 6) + 
  theme_classic() + 
  geom_smooth(method = lm, color = "black") +
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  labs(color = "Mother's ID") + 
  xlab("Log Estradiol Concentration (pg/ml)") + 
  ylab("Log Estradiol Concentration (pg/mg)")+
  scale_shape_manual(values = shape_mapping)

momoff_E

quickmodelRe <- summary(lm(mother_offspring_avghormones$Avg_E2.x~mother_offspring_avghormones$Avg_E2.y))

quickmodelRe 
```


```{r}
correlations_together <- momoff_T / momoff_E 
correlations_together

ggsave(filename = "motheroffspring_hormones.png", path = "Pogona Hormones/output", width = 13, height = 15)
```


### MC versions of mother/offspring plots

```{r}

momoff_T <- ggplot(mc_motheroffspring, aes(x = Avg_T, y = Avg_mcT, color = MomID)) + 
  geom_point(aes(shape = MomGeno),size = 6) + 
  theme_classic() + 
  geom_smooth(method = lm, color = "black") +
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  labs(color = "Mother's ID") + 
  xlab("Log Testosterone Concentration (pg/ml)") + 
  ylab("Log Testosterone Concentration (pg/mg)")+
  scale_shape_manual(values = shape_mapping)

momoff_T

quickmodelT <- summary(lm(mc_motheroffspring$Avg_T~mc_motheroffspring$Avg_mcT))

quickmodelT 
```

```{r}

momoff_E <- ggplot(mc_motheroffspring, aes(x = Avg_E2, y = Avg_mcE2, color = MomID)) + 
  geom_point(aes(shape = MomGeno), size = 6) + 
  theme_classic() + 
  geom_smooth(method = lm, color = "black") +
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  labs(color = "Mother's ID") + 
  xlab("Log Estradiol Concentration (pg/ml)") + 
  ylab("Log Estradiol Concentration (pg/mg)")+
  scale_shape_manual(values = shape_mapping)

momoff_E

quickmodelRe <- summary(lm(mc_motheroffspring$Avg_E2~mc_motheroffspring$Avg_mcE2))

quickmodelRe 
```


```{r}
correlations_together <- momoff_T / momoff_E 
correlations_together

ggsave(filename = "motheroffspring_hormones.png", path = "Pogona Hormones/output", width = 13, height = 15)
```


```{r}

momoff_T <- ggplot(mc_motheroffspring, aes(x = Avg_E2, y = Avg_mcT, color = MomID)) + 
  geom_point(aes(shape = MomGeno),size = 6) + 
  theme_classic() + 
  geom_smooth(method = lm, color = "black") +
  theme(text = element_text(size = 25)) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  labs(color = "Mother's ID") + 
  xlab("Log Estradiol Concentration (pg/ml)") + 
  ylab("Log Testosterone Concentration (pg/mg)")+
  scale_shape_manual(values = shape_mapping)

momoff_T

quickmodelT <- summary(lm(mc_motheroffspring$Avg_E2~mc_motheroffspring$Avg_mcT))

quickmodelT 

ggsave(filename = "offT_mothE_hormones.png", path = "Pogona Hormones/output", width = 13, height = 15)
```


### Mean egg sizes of clutch mothers

```{r}
ggplot(yolkdatv3, aes(y = EggMass, x = MomID, fill = MomGeno)) + geom_boxplot() + geom_point() + scale_fill_manual(values = c( "#dab0da","#809063")) + theme_classic() + xlab("Mother's ID") + ylab("Egg Mass (g)") + theme(text = element_text(size = 25)) + theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) + ylim(2.75, 5)

ggsave(filename = "MassbyMother.png", path = "Pogona Hormones/output", width = 10, height = 12)

```


### Models to look at sources of variance in the dataset

```{r}
pacman::p_load(lme4, rptR, lmerTest)
```


### Base models and rptR bootstrapping

```{r}
#Base T model

pog_modelT1<-lmer(logT ~ 1 + (1|MomID), data = yolkdat)

summary(pog_modelT1)
```

```{r}
#Base T model

pog_modelT1<-lmer(logT ~ 1 + (1|MomID), data = yolkdatv3)

summary(pog_modelT1)
```

```{r}
#Bootstrapping repeatibility and uncertainty with rptR
#Quick histo as a normality check, using gaussain distribution for rptR
hist(yolkdatv3$logT)

#Base test rpt no bootstrapping

rpt(logT ~ 1 + (1|MomID), grname = "MomID", data = yolkdatv3, datatype = "Gaussian", 
    nboot = 0, npermut = 0)

```

```{r}

#Bootstrapped repeatibility 

rptT1 <- rpt(logT ~ 1 + (1|MomID), grname = "MomID", data = yolkdatv3, datatype = "Gaussian", 
    nboot = 5000, npermut = 0)

#Pretty massive confidence intervals on the variance explained by Mother/Clutch, but does not cross O and has p<0.005? Have I interpreted this correctly?

```


```{r}
#Base E model

pog_modelE1<-lmer(logE ~ 1 + (1|MomID), data = yolkdatv3)

summary(pog_modelE1)
```

```{r}
#Bootstrapping repeatibility and uncertainty with rptR
#Quick histo as a normality check, using gaussain distribution for rptR
hist(yolkdatv3$logE)

#Base test rpt no bootstrapping

rpt(logE ~ 1 + (1|MomID), grname = "MomID", data = yolkdatv3, datatype = "Gaussian", 
    nboot = 0, npermut = 0)

```

```{r}

#Bootstrapped repeatibility 

rptE1 <- rpt(logE ~ 1 + (1|MomID), grname = "MomID", data = yolkdatv3, datatype = "Gaussian", 
    nboot = 5000, npermut = 0)


```


### Model 2 with MomGeno

```{r}
# T model with MomGeno

pog_modelT2<-lmer(logT ~ MomGeno + (1|MomID), data = yolkdatv3)

summary(pog_modelT2)
```

```{r}

#Bootstrapped repeatibility 

rptT2 <- rpt(logT ~ MomGeno + (1|MomID), grname = "MomID", data = yolkdatv3, datatype = "Gaussian", 
    nboot = 5000, npermut = 0)


```




```{r}
# E model with MomGeno

pog_modelE2<-lmer(logE ~ MomGeno + (1|MomID), data = yolkdatv3)

summary(pog_modelE2)
```

### Manuscript Yolk Figure
```{r}
Tyolksv3 <- ggplot(yolkdatv3, aes(x=MomID, y=logT, fill = MomGeno)) + geom_violin(alpha = 0.75, width = 1) + geom_boxplot(width = 0.05, position = position_dodge(0.9)) + geom_point(size = 5, position = position_dodge(width = 0.2), alpha = 0.75) + scale_fill_manual(values = c("plum3", "darkolivegreen")) + ylab("Log Testosterone Concentration (pg/mg)") + xlab("Mother's ID") + theme_classic() +  theme(text = element_text(size = 25)) + ylim (-2.5, 1) + theme(axis.line = element_line(size = 1), axis.ticks = element_line(size = 1)) +
  theme(axis.ticks.length = unit(0.3, "cm"))

Tyolksv3 


ggsave(filename = "AAA3_TestoClutch_plot.png", path = "Pogona Hormones/output", width = 15, height = 10)
```

```{r}

Eyolks3 <- ggplot(yolkdatv3, aes(x = MomID, y = logE, fill = MomGeno)) +
  geom_violin(alpha = 0.75, width = 1) +
  geom_boxplot(width = 0.05, position = position_dodge(0.9)) +
  geom_point(size = 5, position = position_dodge(width = 0.2), alpha = 0.75) +
  scale_fill_manual(values = c("plum3", "darkolivegreen")) +
  ylab("Log Estradiol Concentration (pg/mg)") +
  xlab("Mother's ID") +
  theme_classic() +
  theme(text = element_text(size = 25)) +
  coord_cartesian(ylim = c(-3.5, 0)) +  # Preserve ylim settings
  theme(axis.line = element_line(size = 1), 
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  scale_y_continuous(labels = function(x) sprintf("%.1f", x)) 

print(Eyolks3)
ggsave(filename = "AAA3_EstradiolClutch_plot.png", path = "Pogona Hormones/output", width = 15, height = 10)

```

```{r}
finalyolkplot3<- Tyolksv3 / Eyolks3

finalyolkplot3

ggsave(filename = "AAA3_Clutch_plot.png", path = "Pogona Hormones/output", width = 16, height = 15)
```

### Manuscript plot retry
```{r}
library(ggplot2)
library(patchwork)  # For combining plots

#Order MomID so that the ZW and ZZ females are presented next to each other

yolkdatv3$MomID <- factor(yolkdatv3$MomID, levels = c("9520", "9555", "9559", "81803", "3505", "9152", "9570", "82142"))

# Plot for Testosterone Concentration
Tyolksv3 <- ggplot(yolkdatv3, aes(x = MomID, y = logT, fill = MomGeno)) +
  geom_violin(alpha = 0.75, width = 0.4, scale = "width") +
  geom_boxplot(width = 0.05, position = position_dodge(0.9)) +
  geom_point(size = 5, position = position_dodge(width = 0.2), alpha = 0.75) +
  scale_fill_manual(values = c("plum3", "darkolivegreen")) +
  ylab("Log Testosterone Concentration (pg/mg)") +
  xlab("Mother's ID") +
  theme_classic() +
  theme(text = element_text(size = 25)) +
  ylim(-2.5, 1) +
  theme(axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm"))

print(Tyolksv3)
ggsave(filename = "AAA3_TestoClutch_plot.png", path = "Pogona Hormones/output", width = 15, height = 10)

# Plot for Estradiol Concentration
Eyolks3 <- ggplot(yolkdatv3, aes(x = MomID, y = logE, fill = MomGeno)) +
  geom_violin(alpha = 0.75, width = 0.4, scale = "width") +
  geom_boxplot(width = 0.05, position = position_dodge(0.9)) +
  geom_point(size = 5, position = position_dodge(width = 0.2), alpha = 0.75) +
  scale_fill_manual(values = c("plum3", "darkolivegreen")) +
  ylab("Log Estradiol Concentration (pg/mg)") +
  xlab("Mother's ID") +
  theme_classic() +
  theme(text = element_text(size = 25)) +
  coord_cartesian(ylim = c(-3.5, 0)) +  # Preserve ylim settings
  theme(axis.line = element_line(size = 1), 
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  scale_y_continuous(labels = function(x) sprintf("%.1f", x))

print(Eyolks3)
ggsave(filename = "AAA3_EstradiolClutch_plot.png", path = "Pogona Hormones/output", width = 15, height = 10)

# Combine the two plots
finalyolkplot3 <- Tyolksv3 / Eyolks3
```

```{r}
print(finalyolkplot3)
ggsave(filename = "anomaloplot.png", path = "Pogona Hormones/output", width = 16, height = 15)
```

### Mean cluthc measurements for ZZ and ZW clutches

```{r}
Eyolksmean <- ggplot(yolkdatv3, aes(x = MomGeno, y = logE, fill = MomGeno)) +
  geom_violin(alpha = 0.75, width = 0.4, scale = "width") +
  geom_boxplot(width = 0.1, position = position_dodge(0.9)) +
  geom_point(size = 5, position = position_dodge(width = 0.2), alpha = 0.75) +
  scale_fill_manual(values = c("plum3", "darkolivegreen")) +
  ylab("Log Estradiol Concentration (pg/mg)") +
  xlab("Mother's Genotype") +
  theme_classic() +
  theme(text = element_text(size = 25)) +
  coord_cartesian(ylim = c(-3.5, 0)) +  # Preserve ylim settings
  theme(axis.line = element_line(size = 1), 
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  scale_y_continuous(labels = function(x) sprintf("%.1f", x))

print(Eyolksmean)

```

```{r}
Tyolksmean <- ggplot(yolkdatv3, aes(x = MomGeno, y = logT, fill = MomGeno)) +
  geom_violin(alpha = 0.75, width = 0.4, scale = "width") +
  geom_boxplot(width = 0.1, position = position_dodge(0.9)) +
  geom_point(size = 5, position = position_dodge(width = 0.2), alpha = 0.75) +
  scale_fill_manual(values = c("plum3", "darkolivegreen")) +
  ylab("Log Testosterone Concentration (pg/mg)") +
  xlab("Mother's Genotype") +
  theme_classic() +
  theme(text = element_text(size = 25)) +
  coord_cartesian(ylim = c(-2.5, 1)) +  # Preserve ylim settings
  theme(axis.line = element_line(size = 1), 
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm")) +
  scale_y_continuous(labels = function(x) sprintf("%.1f", x))

Tyolksmean
```

```{r}
yolksmean <- Tyolksmean / Eyolksmean

yolksmean

ggsave(filename = "AAAyolksmean_genotype.png", path = "Pogona Hormones/output", width = 10, height = 15)
```